### 2022-04-03 Project/ltrsPlots.Rmd
ltrsPlots

---
title: "R Notebook"
output: html_notebook
---

## What is already coded
```{r}
library(ape)
library(tidyr)
library(MASS)
library(MuMIn)
library(ggplot2)
library(reshape2)
library(grid)
library(gridExtra)
library(ggfortify)
library(vegan)
library(ggtree)
library(dplyr)

###NK & HB lifetime reproductive success data
ltrsHB <- read.csv("./Data/HargreavesEcolLett data ltrsHB.csv")
ltrsNK <- read.csv("./Data/HargreavesEcolLett data ltrsNK.csv")

```

```{r}
# Select variables
ltrsHB_clean <- ltrsHB %>% select(year:treat, Source:totbadseed)
ltrsNK_clean <- ltrsNK %>% select(year:treat, Source:tbo, Ln.tot:totbadseed)

# Merge datasets
ltrsdat <- rbind(ltrsHB_clean, ltrsNK_clean)

# Remove data not from High, Medium, or Low elevation 
elevation <- c("H", "M", "L")
ltrsdat <- ltrsdat %>% 
  filter(Source %in% elevation)

# Examine variable classes
str(ltrsdat)

```

## New Code for Plots

Compare Leaf nodes between treatments
```{r}

# Create plot data
ltrsPlotDat <- ltrsdat %>% mutate(propgoodseed = totgoodseed/(totbadseed + totgoodseed)) %>% 
  select(Source, site, treat, Ln.tot, Fr.tot, propgoodseed)

# Reorder sites
ltrsPlotDat$site <- factor(ltrsPlotDat$site , levels = c("L", "M", "H", "HE", "A", "A1", "A2"))

# Filter for mid and high elevations
MidPlotDat <- ltrsPlotDat %>%  filter(Source == "M")
HighPlotDat <- ltrsPlotDat %>%  filter(Source == "H")

# Mid Elevation Source Plot
# Create Mid Elevation plot
MidLeafPlot <- ggplot(MidPlotDat, aes(x = site, y = Ln.tot, fill = treat)) +
  geom_boxplot(width = 0.5) +
  scale_x_discrete(name = "\n Transplant site", 
                   labels = c("Low elevation", "Mid-elevation", "High elevation", 
                              "High elevation range edge", "Above range HB",
                              "First site above range NK", "Second site above range NK")) +
  scale_y_continuous(name = "Total leaf nodes \n")+
  scale_fill_discrete(name = "Treatment", labels = c("Warmed", "Control")) +
  ggtitle("Leaf growth of mid-elevation sourced plants") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
        plot.title = element_text(hjust = 0.5))

MidLeafPlot

# High Elevation Source Plot
# Create Mid Elevation plot
HighLeafPlot <- ggplot(HighPlotDat, aes(x = site, y = Ln.tot, fill = treat)) +
  geom_boxplot(width = 0.5) +
  scale_x_discrete(name = "\n Transplant site", 
                   labels = c("Low elevation", "Mid-elevation", "High elevation", 
                              "High elevation range edge", "Above range HB",
                              "First site above range NK", "Second site above range NK")) +
  scale_y_continuous(name = "Total leaf nodes \n")+
  scale_fill_discrete(name = "Treatment", 
                      labels = c("Warmed", "Control")) +
  ggtitle("Leaf growth of high elevation sourced plants") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
        plot.title = element_text(hjust = 0.5))

HighLeafPlot

```

Compare fruits between treatments
```{r}

# Mid Elevation Source Plot
# Create Mid Elevation plot
MidFruitPlot <- ggplot(MidPlotDat, aes(x = site, y = Fr.tot, fill = treat)) +
  geom_boxplot(width = 0.5) +
  scale_x_discrete(name = "\n Transplant site", 
                   labels = c("Low elevation", "Mid-elevation", "High elevation", 
                              "High elevation range edge", "Above range HB",
                              "First site above range NK", "Second site above range NK")) +
  scale_y_continuous(name = "Total fruits from all stems \n")+
  scale_fill_discrete(name = "Treatment", labels = c("Warmed", "Control")) +
  ggtitle("Fruit development of mid-elevation sourced plants") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
        plot.title = element_text(hjust = 0.5))

MidFruitPlot

# High Elevation Source Plot
# Create Mid Elevation plot
HighFruitPlot <- ggplot(HighPlotDat, aes(x = site, y = Fr.tot, fill = treat)) +
  geom_boxplot(width = 0.5) +
  scale_x_discrete(name = "\n Transplant site", 
                   labels = c("Low elevation", "Mid-elevation", "High elevation", 
                              "High elevation range edge", "Above range HB",
                              "First site above range NK", "Second site above range NK")) +
  scale_y_continuous(name = "Total fruits from all stems \n")+
  scale_fill_discrete(name = "Treatment", 
                      labels = c("Warmed", "Control")) +
  ggtitle("Fruit development of high elevation sourced plants") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
        plot.title = element_text(hjust = 0.5))

HighFruitPlot

```

Compare total good seeds between sites for each source
```{r}

# Mid Elevation Source Plot
# Create Mid Elevation plot
MidSeedPlot <- ggplot(MidPlotDat, aes(x = site, y = propgoodseed, fill = treat)) +
  geom_boxplot(width = 0.5) +
  scale_x_discrete(name = "\n Transplant site", 
                   labels = c("Low elevation", "Mid-elevation", "High elevation", 
                              "High elevation range edge", "Above range HB",
                              "First site above range NK", "Second site above range NK")) +
  scale_y_continuous(labels = scales::percent, 
                     name = "Percentage of viable seeds produced by plant \n")+
  scale_fill_discrete(name = "Treatment", labels = c("Warmed", "Control")) +
  ggtitle("Seed viability of mid-elevation sourced plants") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
        plot.title = element_text(hjust = 0.5))

MidSeedPlot

# High Elevation Source Plot
# Create Mid Elevation plot
HighSeedPlot <- ggplot(HighPlotDat, aes(x = site, y = propgoodseed, fill = treat)) +
  geom_boxplot(width = 0.5) +
  scale_x_discrete(name = "\n Transplant site", 
                   labels = c("Low elevation", "Mid-elevation", "High elevation", 
                              "High elevation range edge", "Above range HB",
                              "First site above range NK", "Second site above range NK")) +
  scale_y_continuous(labels = scales::percent,
                     name = "Percentage of viable seeds produced by plant \n")+
  scale_fill_discrete(name = "Treatment", 
                      labels = c("Warmed", "Control")) +
  ggtitle("Seed viability of high elevation sourced plants") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 7),
        plot.title = element_text(hjust = 0.5))

HighSeedPlot

```